\documentclass[11pt]{article}
\usepackage{times}
\usepackage{fullpage}
\usepackage{url}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{amsmath, amsfonts, amsthm, fouriernc}
\usepackage{IEEEtrantools}
\usepackage{parskip}

\pagestyle{fancy}
\addtolength{\headheight}{2em}
\addtolength{\headsep}{1.5em}
\lhead{\large{ME 2801}}

\usepackage{sectsty}
\sectionfont{\fontsize{12}{8}\selectfont}

\begin{document}

\begin{center}
  \Large{\bf{Assignment: Steady-State Error}}
\end{center}

The textbook exercises, designated with the word ``Nise'', are from the
``Problems'' section at the end of the chapter in \emph{Control System
Engineering} by Norman Nise, 7th edition.

% ---------------------------------------------------------------
\section*{Nise 7.5}

Given the open-loop transfer function, predict the steady-state error for step
(position), ramp (velocity), and parabolic (acceleration) test inputs.  By
inspection of $G(s)$ you should be able to anticipate the type of steady-state
error (zero, constant, or infinite) for each input type before computing it.

% ---------------------------------------------------------------
\section*{Nise 7.10}

Use the \emph{Key Equations} (inside back cover of the Nise textbook) to find
the answers by hand.  Optionally, check with MATLAB.

% ---------------------------------------------------------------
\section*{Nise 7.17}

This exercise revisits the derivation of the steady-state error equations for
step, ramp, and parabolic inputs to deepen understanding of system type.

% ---------------------------------------------------------------
\section*{USV Case Study}

From the previous assignment, you developed proportional-only unity-feedback
controllers for surge (speed) and yaw, and evaluated the closed-loop response
to step inputs.  You may have observed that the two controllers had different
steady-state error behavior.  Here you will use system-type analysis to explain
that difference.

\subsection*{Plant models}

Surge (speed) plant:
\[
  G_u(s) = \frac{1}{Ms + D}
\]
Proportional feedback gain: $K_u$.

Yaw (heading) plant:
\[
  G_r(s) = \frac{1}{s\,(Is + d)}
\]
Proportional feedback gain: $K_r$.

\subsection*{Exercises}

Work through the following steps independently for each degree of freedom
(surge and yaw).

\begin{itemize}

  \item What is the system Type for each closed loop?

  \item Based on the system Type, which input (step, ramp, or parabolic)
    produces a non-zero, finite steady-state error?

  \item For that input type, find an expression for the constant steady-state
    error in terms of the relevant physical and control parameters
    ($K_u$, $K_r$, $M$, $D$, $I$, $d$).
    \begin{itemize}
      \item How do the physical model parameters affect (increase or decrease)
        steady-state error?
      \item How does the control gain affect steady-state error?
    \end{itemize}

\end{itemize}

\end{document}
