\documentclass[11pt]{article}
\usepackage{times}
\usepackage{fullpage}
\usepackage{url}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{amsmath, amsfonts, amsthm, fouriernc}
\usepackage{IEEEtrantools}
\usepackage{parskip}
\usepackage{booktabs}

\pagestyle{fancy}
\addtolength{\headheight}{2em}
\addtolength{\headsep}{1.5em}
\lhead{\large{ME 2801}}
\rhead{\large{Solutions}}

\usepackage{sectsty}
\sectionfont{\fontsize{12}{8}\selectfont}

% Auto-numbered exercise headings: \exercise{Title} -> "Exercise N --- Title"
\newcounter{exercise}
\newcommand{\exercise}[1]{%
  \refstepcounter{exercise}%
  \section*{Exercise \theexercise\ --- #1}%
}

\begin{document}

\begin{center}
  \Large{\bf{Solutions: Steady-State Error}}
\end{center}

Numerical verification for every exercise is in
\texttt{code/assignment\_solutions.m}.  Run section-by-section.

% ---------------------------------------------------------------
\exercise{Three Test Inputs}

\textbf{(a)}  $G(s)$ has no factors of $s$ in the denominator
$\Rightarrow$ \textbf{Type 0}.

\textbf{(b)}  Steady-state errors follow from the master equation
\[
e(\infty) = \lim_{s \to 0} s \frac{R(s)}{(1+G(s))},
\] with each input
$R(s) = 1/s,\ 1/s^2,\ 1/s^3$ in turn.  The relevant DC limits of $G(s)$ are
\[
  K_x \equiv \lim_{s \to 0} G(s) = \frac{600}{(10)(25)} = 2.4,
  \qquad
  \lim_{s \to 0} s\,G(s) = 0,
  \qquad
  \lim_{s \to 0} s^2\,G(s) = 0,
\]
giving
\[
  e_\mathrm{step} = \frac{1}{1 + K_x} = \frac{1}{3.4} \approx \boxed{0.294},
  \qquad
  e_\mathrm{ramp} = \boxed{\infty},
  \qquad
  e_\mathrm{para} = \boxed{\infty}.
\]

\textbf{(c)}  In MATLAB:
\begin{itemize}
  \item \emph{Calculate.}  \texttt{T = feedback(G,1); 1 - dcgain(T)}
    returns 0.2941.
  \item \emph{Simulate.}  \texttt{step(T)} settles at
    $0.706 = 1 - 0.294$, so the error signal $r - y$ settles at $0.294$.
\end{itemize}
Both match the analytical answer.

\textbf{(d)}  $G(s)$ contains \emph{no integrators} (no factor of $s$ in the
denominator).  Without integration in the forward path, the controller
cannot accumulate the unbounded bias required to keep up with an unbounded
reference; the error grows without limit for the ramp and parabola.

% ---------------------------------------------------------------
\exercise{Design a P-Only Velocity Tracking Compensator}

\textbf{(a)}  The combined open-loop (forward path) is
$G(s) = C(s)\,G_p(s) = \frac{K_p\,(s+5)}{s\,(s+8)(s+12)}$, % CLAUDE: Write as standalone and use \frac
which has one factor of $s$
in the denominator $\Rightarrow$ \textbf{Type 1}.

\textbf{(b)}  Velocity constant: % CLAUDE:  Don't use the velcity constant.   Just use the master formula for the ramp error.
\[
  K_v = \lim_{s \to 0} s\,C(s)\,G_p(s)
      = \lim_{s \to 0} \frac{K_p\,s\,(s+5)}{s\,(s+8)(s+12)}
      = \frac{5\,K_p}{96}.
\]
A 4\,\% ramp error requires $1/K_v = 0.04$, i.e.\ $K_v = 25$:
\[
  \frac{5\,K_p}{96} = 25
  \;\Longrightarrow\;
  \boxed{K_p = 480}.
\]

\textbf{(c)}  The closed-loop characteristic polynomial is
\[
  \Delta(s) = s\,(s+8)(s+12) + 480\,(s+5)
            = s^3 + 20\,s^2 + 576\,s + 2400.
\]
\texttt{pole(feedback(Kp*Gp, 1))} with $K_p = 480$ returns roots near
$-4.46 \pm 14.2j$ and $-11.1$ --- all in the left half-plane, so the
closed loop is stable.  Because the chapter's general SSE formula assumes
stability of the closed loop, the result in (b) is meaningful.

% CLAUDE: instead just say "Use MATLAB to compute the closed-loop poles for your value of Kp.  Are they all in the left half-plane?" 

% ---------------------------------------------------------------
\exercise{USV Surge: Predict Before You Simulate}

Let $C(s)$ denote the controller (P or PI), $G_p(s) = 5.1/(1.3s+1)$ the
plant, and $K_{\!f\!f} = 1/G_p(0) = 1/5.1$ the feedforward gain.  With
feedforward the closed-loop transfer function is
\[
  T(s) = \frac{G_p(s)\bigl(C(s) + K_{\!f\!f}\bigr)}{1 + C(s)\,G_p(s)},
\]
and the step error is $e_\mathrm{ss} = 1 - T(0)$.

\textbf{(a) P only.}  $C(s) = K_P = 0.5$.  The forward TF $K_P\,G_p(s)$ is
Type~0 with $K_x = K_P \cdot G_p(0) = 0.5 \cdot 5.1 = 2.55$:
\[
  e_\mathrm{ss} = \frac{1}{1 + K_x} = \frac{1}{3.55} \approx \boxed{0.282\;\mathrm{m/s}}.
\]

\textbf{(b) PI.}  Adding $K_I/s$ to the controller introduces an integrator
$\Rightarrow$ Type~1, so
\[
  e_\mathrm{ss} = \boxed{0}.
\]

\textbf{(c) P + FF.}  $C(0) = 0.5$, so
\[
  T(0) = \frac{5.1\,(0.5 + 1/5.1)}{1 + 5.1 \cdot 0.5}
       = \frac{5.1 \cdot 0.696}{3.55}
       = \frac{3.55}{3.55}
       = 1
  \;\Longrightarrow\;
  e_\mathrm{ss} = \boxed{0}.
\]

\textbf{(d) PI + FF.}  $C(s) \to \infty$ as $s \to 0$ (the $K_I/s$ term
dominates), so $T(0) = 1$ as in (c) and
\[
  e_\mathrm{ss} = \boxed{0}.
\]

\textbf{(e)}  At steady state in a stable closed loop, every signal is
constant, so $\dot{u}_I(t) = 0$.  But $\dot{u}_I = K_I\,e(t)$, so
$\dot{u}_I = 0$ forces $e = 0$ (assuming $K_I \neq 0$).  The integrator
\emph{adapts} until the error has been driven to zero --- it does not need
to know $K_\mathrm{plant}$ to do so.

% ---------------------------------------------------------------
\exercise{USV Yaw: Ramp Tracking Design}

\textbf{(a)}  Heading is the integral of yaw rate, so the heading plant
picks up an extra factor of $1/s$ relative to the yaw-rate plant:
\[
  G_r(s) = \frac{1}{s}\,G_\mathrm{yaw\text{-}rate}(s)
         = \frac{0.82}{s\,(0.61\,s + 1)}.
\]

\textbf{(b)}  $G_r(s)$ has one factor of $s$ in the denominator, and
$C(s) = K_r$ is just a scalar gain.  The combined open-loop $K_r\,G_r(s)$
has the same number of integrators as $G_r$ alone $\Rightarrow$
\textbf{Type 1}.

\textbf{(c)}  Velocity constant:
\[
  K_v = \lim_{s \to 0} s\,K_r\,G_r(s)
      = \lim_{s \to 0} \frac{0.82\,K_r}{0.61\,s + 1}
      = 0.82\,K_r.
\]
For a ramp slope of $5^\circ$/s, the steady-state heading error is
\[
  e_\mathrm{ss} = \frac{5}{K_v} = \frac{5}{0.82\,K_r}\;\;[\text{degrees}].
\]

\textbf{(d)}  Spec: $|e_\mathrm{ss}| \le 1^\circ$:
\[
  \frac{5}{0.82\,K_r} \le 1
  \;\Longrightarrow\;
  K_r \ge \frac{5}{0.82}
        \approx \boxed{6.10}.
\]

\textbf{(e)}  In MATLAB, \texttt{lsim} on the closed loop with $K_r = 6.10$
and a $5t$ reference shows the output settling parallel to the reference
with a constant error of $1.0^\circ$, confirming the spec.

\end{document}
