\documentclass[../../book.tex]{subfiles}
\begin{document}
\graphicspath{{images/}}
\chaptertitle{Lab 2: USV PID Control}{7}

\section{Learning Objectives}

By the end of this lab you should be able to:

\begin{description}
  \item[\textbf{Describe}] How ArduRover implements closed-loop PID speed and steering-rate control, and how it connects to the first-order plant models identified in Lab~1.
  \item[\textbf{Configure}] ArduRover PID parameters to implement a specific control design on the USV autopilot.
  \item[\textbf{Explain}] The role of command input filtering and why disabling the acceleration limit allows a step input to reach the controller directly.
  \item[\textbf{Measure}] Closed-loop step-response metrics---rise time, percent overshoot, settling time, and steady-state error---from field data.
  \item[\textbf{Tune}] Speed and steering-rate PID gains to improve closed-loop performance relative to factory defaults.
  \item[\textbf{Compare}] Closed-loop step-response predictions from the Lab~1 plant model against measured field data for both channels, and interpret where the model guides sound design decisions even when it does not perfectly match the experiment.
  \item[\textbf{Argue}] For a specific tuning tradeoff in the context of a timed waypoint race.
\end{description}

% ============================================================
\section{Introduction: ArduPilot for USVs}

ArduPilot is a mature, open-source autopilot software project supporting a wide range of unmanned platforms---fixed-wing aircraft, multirotors, traditional helicopters, rovers, and surface vessels.  The project grew from aerial roots, and the multirotor and fixed-wing implementations benefit from a large user base, extensive documentation, and years of field-validated defaults.

Applying ArduPilot to an unmanned surface vehicle (USV) is technically straightforward in principle: the same autopilot hardware and the same closed-loop PID architecture that steers a quadrotor can steer a boat.  In practice, however, USV users face a distinct set of challenges:

\begin{itemize}
  \item \textbf{Architectural complexity.}  The ArduRover firmware---ArduPilot's rover and boat implementation---inherits infrastructure designed for a broad class of vehicles.  Much of this complexity is irrelevant to a simple single-screw USV, but it must be understood or carefully bypassed.

  \item \textbf{Aerial-centric documentation.}  Online tutorials, forum threads, and parameter descriptions routinely assume an aerial context.  Terminology such as ``throttle,'' ``heading hold,'' and ``attitude'' carries implicit aerial meaning that must be reinterpreted for a surface vehicle.

  \item \textbf{Small community.}  The USV user base is a small fraction of the ArduPilot project, so there is less empirical data, fewer worked examples, and less community synthesis to guide configuration.

  \item \textbf{Distinct dynamics.}  A single-screw, single-rudder USV is fundamentally different from an overactuated multirotor.  The multirotor is holonomic---it can accelerate in any direction from rest---whereas the USV is non-holonomic: it can only generate lateral force indirectly, by developing forward speed and then deflecting a rudder into the flow.  Additionally, the USV has limited control authority at low speeds and is susceptible to wind and current disturbances that have no direct aerial analogue.
\end{itemize}

These experiments use \textbf{ArduRover V4.6.3 (3fc7011a)}.  Parameter names and defaults may differ in later releases; verify against the ArduRover parameter documentation for the firmware version installed on your vehicle.

% ============================================================
\section{Closed-Loop Control Architecture}

\subsection{ArduRover Operating Mode}

ArduRover supports several control modes.  This lab uses \textbf{ACRO mode}, in which the RC transmitter sticks command \emph{desired} speed and \emph{desired} yaw rate rather than directly commanding throttle and rudder deflection.  ArduRover's PID controllers then close the loop, driving the actuators to track the commanded values.

\begin{center}
\begin{tabular}{L{3.0cm} L{9.5cm}}
  \textbf{Throttle stick} & Commands desired surge speed $r_u$ [m/s] \\
  \textbf{Steering stick} & Commands desired yaw rate $r_r$ [rad/s] \\
\end{tabular}
\end{center}

\vspace{4pt}

In Manual mode the PID controllers are bypassed and the sticks command the actuators directly; the closed-loop behavior described in this lab is not present in Manual mode.

\verify{Confirm ACRO mode stick behavior in ArduRover V4.6.3.}

\subsection{Surge Speed Control Loop}
\label{sec:surge_loop}

The pilot commands a desired speed $r_u(t)$.  This command passes through an optional acceleration limiter (Section~\ref{sec:filtering}) before reaching the speed PID controller.  The PID computes a speed error $e_u = r_u - u$ and produces a normalized throttle command $\delta_T \in [-1, +1]$, which drives the ESC and propeller.  Surge speed $u$ is estimated by ArduRover's Extended Kalman Filter (EKF), which fuses GPS velocity and IMU measurements to produce a higher-rate, lower-noise velocity estimate than raw GPS alone, and fed back to close the loop.

\subsection{Yaw-Rate Control Loop}
\label{sec:yaw_loop}

The steering stick commands a desired yaw rate $r_r(t)$.  The yaw-rate PID computes an error $e_r = r_r - r$ and outputs a normalized rudder command $\delta_R \in [-1, +1]$.  Yaw rate $r$ is measured directly by the IMU's $z$-axis gyroscope and fed back.

\subsection{Connection to Lab~1 Plant Models}

The plant transfer functions $G_{\mathrm{surge}}^*$ and $G_r^*$ are exactly the models you identified in Lab~1.  For any set of PID gains, you can use MATLAB to simulate the closed-loop step response and predict performance metrics before going to the field.

Use this capability deliberately.  Before Run~2, simulate the closed-loop response for your chosen gains and compare it against the Run~1 data.  After the field session, overlay the predicted and measured responses on the same axes.

The linear model is a design tool, not a truth model.  The Lab~1 identification invested limited time and used simplified dynamics---linear drag, decoupled surge and yaw channels, no wind or current.  The purpose of the model is to guide design choices: to tell you whether increasing $K_p$ will likely shorten rise time, or whether adding integral action will eliminate steady-state error.  Expect qualitative agreement and treat discrepancies as information about what the model misses.

% ============================================================
\section{Command Input Filtering}
\label{sec:filtering}

\subsection{The Acceleration Limiter in ArduRover}

ArduRover applies a nonlinear rate limiter to the speed command before it reaches the PID controller.  As covered in the Week~5 input-filtering companion script, a rate limiter ramps the command at a fixed maximum rate rather than stepping it instantaneously.  For speed commands, the limiting rate is set by the parameter \texttt{ATC\_ACCEL\_MAX} [m/s$^2$].

With a non-zero \texttt{ATC\_ACCEL\_MAX}, moving the throttle stick instantly produces a setpoint that \emph{ramps} at $R_{\max}$ m/s$^2$.  The PID never sees a true step; it sees a ramp until the setpoint reaches the commanded value.  This is the same rate-limiter behavior studied in Week~5, and it has two practical effects on the step-response experiment:

\begin{itemize}
  \item The effective input to the closed-loop system is a ramp, not a step, so rise-time and overshoot measurements reflect both the limiter and the PID dynamics together.
  \item A faster rise time can be ``achieved'' simply by increasing $R_{\max}$---without any change to the PID gains.  This conflates command rate-limiting with controller performance.
\end{itemize}

\subsection{Disabling the Limiter for Testing}

To observe the PID's true closed-loop step response, set:

\begin{center}
  \texttt{ATC\_ACCEL\_MAX} $= 0$ \quad (unlimited; speed setpoint steps immediately with the stick)
\end{center}

With this setting, a quick stick movement delivers an approximate step directly to the PID error input.  The closed-loop step response then reflects only the PID gains and the plant dynamics---consistent with the analysis in Sections~\ref{sec:surge_loop} and~\ref{sec:yaw_loop}.

Note: In a fielded autonomous system, the acceleration limiter protects actuators and preserves stability margin.  Disable it only for the manual step-response characterization in this lab.

% ============================================================
\section{ArduRover Parameter Configuration}
\label{sec:params}

\subsection{Category 1: Preset Parameters}

The following parameters are configured by the instructor before the lab.  The \textbf{Default} column shows values from the vehicle parameter file; \textbf{Lab Setting} shows the value to use for the step-response experiments.  Bold lab settings indicate a change from the default.

\vspace{6pt}
\begin{center}
\begin{tabular}{L{3.6cm} C{1.8cm} C{1.8cm} L{5.8cm}}
  \textbf{Parameter} & \textbf{Default} & \textbf{Lab Setting} & \textbf{Rationale} \\
  \midrule
  \texttt{ATC\_ACCEL\_MAX}     & 1     & \textbf{0} & Disables speed command rate limiter (Section~\ref{sec:filtering}); allows a true step to reach the speed PID \\[6pt]
  \texttt{ATC\_STR\_ACC\_MAX}  & 120   & \textbf{0} & Disables steering rate acceleration limiter; same purpose for the yaw-rate channel \\[6pt]
  \texttt{ATC\_STR\_RAT\_FF}   & 2.0   & \textbf{0} & Removes feedforward so the step response reflects PID gains only; restore after the lab \\[6pt]
  \texttt{ATC\_SPEED\_IMAX}    & 1.0   & 1.0        & Integrator anti-windup limit; appropriate for this vehicle \\[6pt]
  \texttt{ATC\_STR\_RAT\_IMAX} & 1.0   & 1.0        & Steering-rate integrator anti-windup limit; appropriate for this vehicle \\[6pt]
  \texttt{ATC\_SPEED\_FLTE}    & 10 Hz & 10 Hz      & Low-pass filter on the speed error signal; already active \\[6pt]
  \texttt{ATC\_STR\_RAT\_FLTE} & 10 Hz & 10 Hz      & Low-pass filter on the steering-rate error signal; already active \\[4pt]
  \bottomrule
\end{tabular}
\end{center}
\vspace{6pt}

\subsection{Category 2: Tunable Parameters}

The following parameters are the student's tuning variables.  Default values are from the vehicle parameter file (\texttt{usv\_ardu\_13Apr2026.param}).

\vspace{6pt}
\begin{center}
\begin{tabular}{L{4.0cm} C{2.0cm} L{6.2cm}}
  \textbf{Parameter} & \textbf{Default} & \textbf{Description} \\
  \midrule
  \texttt{ATC\_SPEED\_P}       & 1.0 & Surge speed proportional gain $K_{p,u}$ \\[2pt]
  \texttt{ATC\_SPEED\_I}       & 0.2 & Surge speed integral gain $K_{i,u}$ \\[2pt]
  \texttt{ATC\_SPEED\_D}       & 0.0 & Surge speed derivative gain $K_{d,u}$ \\[6pt]
  \texttt{ATC\_STR\_RAT\_P}    & 0.2 & Steering rate proportional gain $K_{p,r}$ \\[2pt]
  \texttt{ATC\_STR\_RAT\_I}    & 0.2 & Steering rate integral gain $K_{i,r}$ \\[2pt]
  \texttt{ATC\_STR\_RAT\_D}    & 0.0 & Steering rate derivative gain $K_{d,r}$ \\[4pt]
  \bottomrule
\end{tabular}
\end{center}
\vspace{6pt}

% ============================================================
\section{Measured Performance Metrics}

For each step-response trial we will extract rise time $t_r$, percent overshoot \%OS, settling time $t_s$, and steady-state error $e_{ss}$ from the data using MATLAB (definitions in the Week~4 time-response chapter).  We will also record \textbf{control effort}---the actuator command ($\delta_T$ for surge, $\delta_R$ for yaw rate)---which reflects how hard the controller is working and places a real physical bound on how aggressive a set of PID gains can be in practice.

% ============================================================
\section{Experiment Protocol}
\label{sec:protocol}

\subsection{Pre-Experiment Setup}

Assign the same roles as Lab~1 (pilot, planner, scribe).  Dockside, before launching:

\begin{enumerate}
  \item Connect to the vehicle via Mission Planner and open the Full Parameter List.
  \item Verify the preset parameters from Section~\ref{sec:params}; the instructor may configure these in advance.
  \item Record the tunable PID gains (Section~\ref{sec:params}) in the field notes \emph{and} save a timestamped or run-labelled export of the full parameter file from Mission Planner.  The handwritten record is for at-a-glance reference during the session; the exported \texttt{.param} file is the authoritative copy that can be reloaded later.
  \item Verify \texttt{LOG\_BITMASK = 65535} so that all ArduRover data is logged to the SD card.
  \item Confirm that the arming switch is wired to restart logging.  Each arming event begins a fresh \texttt{.BIN} log on the SD card.  The scribe is responsible for recording the wall-clock arming time and listing the experiments performed in each log so that runs can be matched to logs during analysis.
\end{enumerate}

The boat enters the water in Manual mode and is switched to \textbf{ACRO mode} only after it is clear of the launch and pointed at open water.  ACRO is the mode used for every step-response trial in this lab; Manual is used to maneuver to the test area and to recover at the end of a run.

\subsection{Run 1: Factory Defaults}

With the tunable parameters at factory defaults:

\begin{enumerate}
  \item Surge step trials (2--3 successful runs):
    \begin{itemize}
      \item Confirm the vehicle is in ACRO mode and arm to start a fresh log; the scribe records the arming time and notes that this log contains the surge step trials.
      \item Begin with the vehicle at rest on open water.
      \item Apply a step throttle command ($\delta_0 \approx 20$--$50\%$ of full deflection, consistent with Lab~1) and hold it until the speed reaches a clear steady state.
      \item The scribe records the time and step fraction for each trial.
      \item Disarm at the end of the surge trials to close the log.
    \end{itemize}

  \item Yaw-rate step trials (2--3 successful runs):
    \begin{itemize}
      \item Re-arm to open a new log file for the yaw-rate trials; the scribe records the arming time and notes that this log contains the yaw-rate step trials.
      \item Begin with the vehicle moving at a low, approximately constant surge speed (to give the rudder flow authority).
      \item Apply a step steering command and hold until the yaw rate settles.
      \item Record the surge speed, step fraction, and direction (positive/negative) for each trial.
    \end{itemize}

  \item Do not adjust any parameters during Run~1.
\end{enumerate}


\subsection{Tune Parameters}
\label{sec:tune}

Bring the Run~1 logs onto a laptop between runs.  Plot the surge and yaw-rate step responses with the actuator effort below each, and read off rise time, percent overshoot, settling time, and steady-state error.  This is the diagnostic input for tuning.

The tuning process is \emph{ad hoc}: there is no single procedure that produces the right gains, but the symptom-to-action mapping at the end of the PID Control chapter (Section~\ref{sec:tuning}, with the troubleshooting summary in Table~\ref{tab:troubleshoot}) gives a starting point.  Apply it as follows:

\begin{enumerate}
  \item Decide which symptom matters most given the race objective.  In a timed waypoint race the figure of merit is total time around the course (Section~\ref{sec:tradeoffs}).  Sluggish rise time and steady-state speed error directly cost time.  Modest surge overshoot is usually tolerable.  Yaw-rate overshoot causes heading hunting and adds traveled distance, also a time cost.  Pick the symptom whose elimination would buy the most time on this course.
  \item Pick one gain to change in one channel.  Use the troubleshooting table: flat steady-state error with a clean transient $\Rightarrow$ raise $K_i$; sluggish rise with no oscillation $\Rightarrow$ raise $K_p$; ringing or sustained oscillation $\Rightarrow$ lower $K_i$ first, then $K_p$.  Change one parameter at a time so that the effect is attributable.
  \item Simulate the candidate gains in MATLAB using your Lab~1 plant model before applying them to the vehicle.  This is free and rules out unstable choices.
  \item Apply the new gain values in Mission Planner, re-arm the vehicle to open a fresh log, and repeat the affected step trial.  The scribe records the new gain values alongside the arming time so that each log can be matched to its parameter set during analysis.
  \item Iterate.  Within a single field session expect two to four iterations per channel rather than a precise optimum.  Stop when the dominant symptom from step~1 is resolved or the next change produces no measurable improvement.
\end{enumerate}

\subsection{Run 2: Tuned Parameters}

After tuning each channel, capture a clean set of step trials with the final gain values.

\begin{enumerate}
  \item Confirm the final gain values in Mission Planner and save a fresh timestamped parameter export labelled ``Run~2.''  Record the same gains in the field notes alongside the reasoning from Section~\ref{sec:tune}.
  \item Re-arm to open a clean log file and repeat the surge step trials at the same step fraction $\delta_0$ used in Run~1, so that the metrics are directly comparable.
  \item Re-arm again for the yaw-rate trials and repeat with the same step fraction used in Run~1.
  \item If time and battery life permit, iterate further on the gains and capture additional clean runs; the scribe labels each new log with its parameter set.
\end{enumerate}

\subsection{Practical Notes}

\begin{itemize}
  \item Keep the surge and yaw experiments separate---do not attempt to measure both channels simultaneously.
  \item Use the same step fraction ($\delta_0$) across Run~1 and Run~2 so that the metrics are directly comparable.
  \item Verify the SD card is logging before each run by checking the log counter in Mission Planner.
\end{itemize}

% ============================================================
\section{Data Analysis}

\subsection{Relevant Log Channels}

Use the same \texttt{bin2mat.py} pipeline as Lab~1 to convert the \texttt{.BIN} log to \texttt{.mat}.  The closed-loop analysis requires the speed and yaw-rate \emph{setpoints} in addition to the measured signals from Lab~1.  These appear in ArduRover's control-tuning log message:

\vspace{6pt}
\begin{center}
\begin{tabular}{L{2.8cm} L{2.8cm} C{1.8cm} L{5.0cm}}
  \textbf{Log message} & \textbf{Field} & \textbf{Units} & \textbf{Description} \\
  \midrule
  \texttt{CTUN} & \texttt{DesSpeed}    & m/s   & Speed setpoint (command to PID) \\[3pt]
  \texttt{GPS}  & \texttt{Spd}         & m/s   & Measured surge speed \\[3pt]
  \texttt{RCOU} & \texttt{C3}          & PWM   & Throttle command to ESC (surge control effort) \\[5pt]
  \texttt{CTUN} & \texttt{DesTurnRate} & rad/s & Yaw-rate setpoint (command to PID) \\[3pt]
  \texttt{IMU}  & \texttt{GyrZ}        & rad/s & Measured yaw rate \\[3pt]
  \texttt{RCOU} & \texttt{C1}          & PWM   & Rudder command to servo (yaw control effort) \\[3pt]
  \bottomrule
\end{tabular}
\end{center}
\vspace{4pt}
{\small Verify \texttt{CTUN} field names against your log using Mission Planner's log browser before scripting the analysis; field names may differ across ArduRover versions.}
\vspace{6pt}

\subsection{Plots to Produce}

Produce the following for \emph{both} channels and \emph{both} runs.  Plots for the two runs should use identical axes and time scales to allow direct comparison.

\subsubsection{Closed-Loop Step Response}

A two-panel figure:
\begin{itemize}
  \item \textbf{Top}: speed setpoint $r_u(t)$ and measured speed $u(t)$ on the same axes.  Annotate $t_r$, \%OS, $y_\infty$, and $e_{ss}$.
  \item \textbf{Bottom}: surge control effort $\delta_T(t)$ (throttle command converted from PWM to $[-1,\,+1]$).
\end{itemize}
Produce an equivalent figure for the yaw-rate channel.

\subsubsection{Model vs.\ Experiment}

For each channel, overlay the predicted closed-loop step response---simulated in MATLAB from the Lab~1 plant model and your chosen PID gains---on the measured response from Run~2.  Annotate any qualitative differences: does the model over- or under-predict rise time?  Does it predict overshoot that isn't observed in the data, or miss overshoot that is?  A brief written interpretation of the overlay is more valuable than a perfect match.

\subsubsection{Performance Summary}

Record the measured metrics in the following table:

\vspace{6pt}
\begin{center}
\begin{tabular}{C{1.8cm} C{1.8cm} C{1.8cm} C{1.8cm} C{1.8cm} C{1.8cm}}
  \textbf{Channel} & \textbf{Run} & \textbf{$t_r$ [s]} & \textbf{\%OS} & \textbf{$t_s$ [s]} & \textbf{$e_{ss}$} \\
  \midrule
  Surge    & Default & & & & \\[6pt]
  Surge    & Tuned   & & & & \\[6pt]
  Yaw rate & Default & & & & \\[6pt]
  Yaw rate & Tuned   & & & & \\[6pt]
  \bottomrule
\end{tabular}
\end{center}
\vspace{6pt}

% ============================================================
\section{Performance Tradeoffs and Tuning Strategy}
\label{sec:tradeoffs}

\subsection{Race Context}

Lab~3 is a timed waypoint race: the USV must navigate a series of waypoints in the shortest total time.  Optimal tuning for this task is not the same as optimal tuning for a general regulation problem.

\begin{itemize}
  \item \textbf{Surge speed.}  Slow acceleration wastes time between waypoints.  Speed overshoot is generally tolerable (the vessel is moving forward), but very large overshoot wastes energy and complicates steering at high speed.  Steady-state speed error means the vessel runs slower than commanded---a direct time penalty.

  \item \textbf{Yaw rate.}  Waypoints require turns.  A slow yaw-rate response produces wide, slow arcs.  Yaw-rate overshoot causes heading oscillation, which requires corrective steering and adds distance traveled.  Steady-state yaw-rate error means the vessel undershoots its turns.

  \item \textbf{Control effort.}  Aggressive gains (fast response, low error) demand large actuator commands.  Excessive throttle cycling stresses the ESC and propeller; excessive rudder cycling increases drag and can destabilize the yaw response.  These are physical constraints on how aggressive your tuning can be.
\end{itemize}

\subsection{Guiding Questions}

Use these questions to frame your team's tuning argument:

\begin{enumerate}
  \item Which metrics matter most for minimizing race time?  Justify with reference to the specific course layout.
  \item What tradeoffs did you accept?  For example, if you increased $K_{p,u}$ to reduce rise time, did overshoot increase?  Is that acceptable given the race context?
  \item Did you include integral action ($K_i > 0$) in one or both channels?  What does the integrator eliminate, and does that matter for the race?
  \item Did you include derivative action ($K_d > 0$)?  Why or why not?
  \item How does the measured closed-loop response compare to what you predicted from the Lab~1 plant model?  Where did the model guide you well, and where did it fall short?
\end{enumerate}

\end{document}
